\textsf{Are large language models superhuman chemists?}

### 2024-10-29 src/tex/appendix.tex
Update src/tex/appendix.tex

\clearpage
\subsection{Tool augmented models} \label{sec:react-environment}
In addition to directly prompting \glspl{llm}, we also investigated the performance of tool-augmented systems.
For this, we investigated the two models which showed the best overall results, \GPTFourO and \ClaudeThreeFiveSonnet (\oone overperformed both models but is not recommended using this model with such \enquote{reasoning} prompts such as ReAct\autocite{yao2022react} or \gls{cot}\autocite{wei2023cot}). 
For both models, we created a ReAct-style tool augmentation environment in which models had access to WolframAlpha, the ArXiv \gls{api}, Wikipedia, and web search (using Brave search \gls{api}).
We based the selection of these tools on the most used tools by humans (see \Cref{fig:tool_use}).
Additionally, we added two specific tools to convert \gls{iupac} names to \gls{smiles}, and \gls{smiles} to \gls{iupac} names.
These conversion tools allow us to better understand how the agents perform for some specific questions in this agent environment configuration.
We implemented the systems using Langchain with the default ReAct prompt and constrained the system to a maximum of ten \gls{llm} calls.

While for \ClaudeThreeFiveSonnet the overall performance increased (by less than 1\%), we observe a decrease in performance for \GPTFourO compared with the \llm without tools (\Cref{tab:performance_table}). 
If we study the results by each type of question, we observe an improvement for questions regarding electron counts or point groups of compounds.
However, the scores decreased for questions asking about the number of isomers or \gls{ghs} pictograms.
For the specific questions about converting IUPAC names to SMILES, the results decreased notably despite the specific tools that were prepared for those questions. 
By studying the reasoning path for these cases, we found an important error in this ReAct configuration resulting from a bad response format of the model that results in an error raised by the LangChain framework.
This show that these agent configurations need to be really optimized for the task, and it is not only about equipping the \gls{llm} with tools, but also it requires of engineering work to improve the results.

@@

>>>>

@@

\clearpage
\subsection{Summary of trends and recommendations}

\oone consistently leads across both the \ChemBench corpus and \chembenchmini.  \ClaudeThreeFiveSonnet, \GPTFourO, \LlamaThreeOneFourZeroFiveBInstruct form the next tier of strong performers. These models show robust performance across all skills and difficulty levels.

Larger models generally perform better (\LlamaThreeOneFourZeroFiveBInstruct > \LlamaThreeOneSeventyBInstruct > \LlamaThreeOneEightBInstruct). However, smaller but well-designed models can compete with larger models. For example, \GemmaTwoNineBIt has an overall accuracy of \variable{output/trends_section_variables/gemma_9B.txt} on the \ChemBench corpus, which is only \variable{output/trends_section_variables/diff_between_llama_405B_and_gemma_9B.txt}\% less accurate than the much larger \LlamaThreeOneFourZeroFiveBInstruct model.

One can observe a clear progression of performance across model families (e.g., \ClaudeThreeFiveSonnet >  \ClaudeThree > \ClaudeTwo or \GPTFourO > \GPTFour > \GPTThreeFiveTurboZeroT). Newer versions consistently outperform their predecessors.
Scores on knowledge-intensive tasks are typically lower than those on calculation and reasoning-intensive questions.

Based on these findings, some scope for improving these models could be to focus on enhancing knowledge-based training, possibly through improved pre-training on chemistry-specific texts or with chemistry knowledge-base integration. However, innovation is also needed to incorporate this information into the systems. For example, even \PaperQATwo could not outperform the \gls{llm} it is using for directing the tools, even though the agent has access to literature evidence.
This suggests we might need to not only focus on building chemistry-specific retrieval datasets since current systems fail to retrieve the relevant papers and should be coupled with more domain-specific databases. However, the observation that our ReAct agents were too fragile to answer questions for which they had custom-made tools available correctly suggests that we also must invest in building more robust agent frameworks (see \Cref{sec:react-environment} for further discussion about the ReAct environment).

Models show the strongest performance in reasoning tasks on the \chembench corpus, which suggests that current training approaches are good at developing logical/reasoning capabilities for basic problem-solving of university-level exams. However, performance drops significantly for advanced tasks, suggesting that more advanced chemistry problems with complex, multi-step solutions (e.g., solving analytical chemistry problems) should be included in training or finetuning.

The post-training (e.g., \gls{rlhf}) of the current models does not equip them with human-like \enquote{chemical taste.} This, however, will be essential for future discovery research (e.g., in combination with genetic algorithms).
